\section{Future Work and Conclusions}\label{sec:conclusions}
In this article, we demonstrated that vision-based localization allows humanoid robots to achieve complex tasks. By integrating the vision algorithm into the control framework, we have been able to gradually
reshape the trajectory in order to compensate for execution errors. This allowed the robot to achieve complex tasks which would be difficult to realize otherwise.

However, several intesresting additions remain to be integrated. First, the local trajectory modifications are not checked to make sure that no auto-collision occurs. Currently, a conservative maximum correction (i.e.\ 2 centimeters\PFA{Again} every two steps) is imposed for safety. By applying recent works such as fast feasibility tests described by~\citet{Perrin10icra}, we would be able to increase the maximum correction without compromising the robot safety. Second, the HRP-2 robot embeds an Inertial Measurement Unit
which could be used to estimate the robot chest attitude. It would definitely be interesting to provide an initial estimation of the robot motion to help the localization process. In addition, we are interested in improving the capabilities of our visual SLAM and vision-based localization systems towards the goal of long-term localization to deal with possible changes in the environment.

\TM{Do we need more future work? / longer conclusion?}
\PFA{No, I think we are fine with that}
%%% Local Variables:
%%% ispell-local-dictionary: "american"
%%% LocalWords:  odometry HRP
%%% End:
